Supplement to Initial estuarine response to the nutrient dense Piney Point wastewater discharge into Tampa Bay, Florida

### 2021-10-20 manu-draft.Rmd
added sg and macro abundance tables and figures

load(file = here('tables/wqsumtab.RData'))
load(file = here('tables/stktab.RData'))
load(file = here('tables/wqcmptab.RData'))
load(file = here('tables/savtab.RData'))
load(file = here('tables/mcrtab.RData'))

@@

Differences in observations between months for water quality, seagrass, and macroalgae within each area (Figure \@ref(fig:map)a) were evaluated using a Kruskal-Wallis one-way analysis of variance (ANOVA) followed by multiple comparisons using 2-sided Mann-Whitney U tests [@Hollander13]. Probability values were adjusted using the sequential Bonferroni method described in [@Holm79] to account for the increased probability of Type I error rates with multiple comparisons. An adjusted p-value < 5\% ($\alpha$ = 0.05) was considered a significant difference between months.  For water quality variables, long-term monthly averages by area were subtracted from 2021 observations to account for normal seasonal variation not attributed to potential effects from Piney Point.  Similar corrections were not done for monthly comparisons of seagrass and macroalgae data because comparable long-term seasonal data do not exist. When appropriate, methods were used to accommodate measured concentrations in water quality variables that were below detection.  These included summary statistics (e.g., median, mean, and standard deviation) following estimates of the empirical cumulative distribution functions for each parameter using the Kaplan-Meier method for censored data [@Helsel05;@Lee20].

@@

From April through September 2021, `r nwq` samples were collected for chl-a, dissolved oxygen, total nitrogen, total phosphorus, total ammonia nitrogen, nitrate/nitrite, pH, salinity, secchi depth, and temperature (Table \@ref(tab:wqsumtab)).  Of these samples, `r outrng`\% were outside of the normal range defined by the long-term monthly monitoring data for the baseline period from 2006 to 2020 (below for Secchi depth, above for all others). The percentage of observations outside of the normal range varied by location and parameter.  For chl-a, `r chlabv[['1']]`\% of the observations were above the normal range for area 1, whereas only `r chlabv[[2]]`\% and `r chlabv[[3]]`\% were above for areas 2 and 3, respectively. Total nitrogen concentrations were above the normal range for `r tnabv[['1']]`\% of observations in area 1, whereas concentrations were above for `r tnabv[['2']]`\% of observations in area 2 and `r tnabv[['3']]`\% in area 3.  Secchi observations were below the normal range for `r secblw[['1']]`\% of observations in area 1 and for `r secblw[['2']]`\% and `r secblw[['3']]`\% of observations in areas 2 and 3.  Notable differences were also observed for dissolved oxygen (e.g., `r dosabv[['1']]`\% were above in area 1, `r dosabv[['2']]`\% in area 2).  Physical parameters (salinity, temperature) were generally within range over the study period. Inorganic nitrogen (ammonia, nitrate/nitrite) was generally within range, although initial time series showed much higher concentrations for ammonia in April near area 1, similar to the effluent measurements in Table \@ref(tab:stktab). Spatial variation among the parameters showed that values were generally above the normal range (or below for Secchi depth) for many locations near Piney Point, Anna Maria Sound, and the northern mouth of Tampa Bay (Figure \@ref(fig:wqmap)).

Total nitrogen, chl-a, and secchi depth followed a temporal progression in 2021 that was distinct from the long-term seasonal trend estimated from historical data (Figure \@ref(fig:wqgamtrnds)).  For area 1, total nitrogen and chlorophyll concentrations were frequently above normal ranges during April.  Concentrations decreased slightly until June and July when values increased again above the seasonal expectation. Many secchi observations in area 1 were lower than normal in April and July.  Observations in areas 2 and 3 were more often within the normal range, with some exceptions for total nitrogen and chl-a in area 3 in  May, June, and July.  Statistical comparisons between months for total nitrogen, chl-a, and secchi depth (Table \@ref(tab:wqcmptab)) supported the results in Figure \@ref(fig:wqgamtrnds). Kruskal-Wallis tests that assessed if at least one of the months had significantly different observations for each parameter were significant for total nitrogen, chl-a, and secchi depth for areas 1 and 3 and total nitrogen and chl-a for area 2 (Table \@ref(tab:wqcmptab)). Results of multiple comparison tests that evaluated differences between pairs of months generally showed that April/May were different from June/July depending on area and parameter.  Observations in the later months were generally higher (or lower for Secchi) corresponding to increasing *K. brevis* abundances.

@@

# fo by month, area
fomo <- trnsum %>% 
  ungroup() %>% 
  mutate(
    foest = round(100 * foest, 1), 
    mo = month(date)
  )
fomoarea1 <- fomo %>% 
  filter(area == 'Area 1')
fomoarea3 <- fomo %>% 
  filter(area == 'Area 3')
fomoarea1apr <- fomoarea1 %>% 
  filter(mo %in% 4) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea1jul <- fomoarea1 %>% 
  filter(mo %in% 7) %>% 
  select(taxa,foest) %>% 
  deframe
fomoarea3apr <- fomoarea3 %>% 
  filter(mo %in% 4) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea3jul <- fomoarea3 %>% 
  filter(mo %in% 7) %>% 
  select(taxa,foest) %>% 
  deframe
fomoarea1jun <- fomoarea1 %>% 
  filter(mo %in% 6) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea3jun <- fomoarea3 %>% 
  filter(mo %in% 6) %>% 
  select(taxa, foest) %>% 
  deframe

# kw pvals
cyanoarea3 <- mcrtab %>% 
  mutate(
    Area = ifelse(Area == '', NA, Area)
  ) %>% 
  fill(Area) %>% 
  filter(Area == 3 & `Macroalgae group` == 'Cyanobacteria') %>% 
  pull(pval) %>% 
  as.numeric %>% 
  round(3)
```

A total of `r ntrns` transects were sampled for seagrasses and macroalgae from April through July, each visited on average `r avevis` times per month. Turtle grass (*T. testudinum*) was the dominant seagrass species with frequency occurrence of `r foall['Thalassia testudinum']`\% across all locations and sample dates.  Manatee grass (*S. filiforme*) and shoal grass (*H. wrightii*) had similar coverage across all transects, with frequency occurrences of `r foall['Syringodium filiforme']`\% and `r foall['Halodule wrightii']`\%, respectively. The frequency occurrences of seagrasses near Piney Point were similar to the long-term record of seagrass transect data available for Tampa Bay [@Sherwood17].  At the baywide scale, shoal grass is the dominant species, whereas turtle grass is more common in euhaline waters closer to the Gulf.  Macroalgae observed along the transects also varied in coverage, with red macroalgae groups having the highest frequency occurrence of `r foall['Red']`\%.  Common taxa in the red group included *Gracilaria spp.* and *Acanthophora sp.*.  Green and cyanobacteria macroalgae were less common, with frequency occurrences of `r foall['Green']`\% and `r foall['Cyanobacteria']`\%.  Common taxa in the green group included *Ulva spp.* and *Caulerpa spp.*, whereas *Dapis spp.* was the only cyanobacteria macroalgae observed.  Brown macroalgae in the genus *Feldmannia* were only observed at `r foall['Brown']`\% frequency occurrence.

Temporal progression of seagrasses and macroalgae varied across the months, although a typical pattern observed at many of the transects is shown in Figure \@ref(fig:trnex).  Transect S3T6 is located less than one kilometer to the north of Port Manatee.  The site is dominated by manatee grass that was observed at nearly all of the sample points along the transect at varying coverages.  Overall abundance of seagrass did not change dramatically from April 22nd through July.  However, macroalgal abundances changed over the course of sampling similar to many of the other transects sampled during the study.  Red macroalgae were present in high abundances from April to May.  *Dapis sp.* was first observed on May 24th and was present at all of the sample locations on June 4th and 15th.  *Dapis sp.* persisted through June and July, but not was observed after July 20th.  Green macroalgae taxa were first observed in July, although at generally low abundances. Trends in macroalgae somewhat followed seasonal patterns observed baywide with red macroalgae transitioning to green macroalgae by mid-summer.

Monthly summaries in frequency occurrence by area (Figure \@ref(fig:trnfrq)) provided an indication of seagrass and macroalgae trends across all transects.  No transects were sampled in area 2 to the north of Piney Point.  Area 1 had generally balanced frequency occurrence of the three primary seagrass species observed in the bay, although a slight decrease over time in shoal grass (*H. wrightii*) was observed (April frequency occurrence of `r fomoarea1apr['Halodule wrightii']`\%, July frequency occurrence of `r fomoarea1jul['Halodule wrightii']`\%).  A similar decline in coverage of shoal grass was also observed in area 3 (April frequency occurrence of `r fomoarea3apr['Halodule wrightii']`\%, July frequency occurrence of `r fomoarea3jul['Halodule wrightii']`\%). Frequency occurrence of manatee grass (*S. syringodium*) also declined slightly from April to July in area 3 (April frequency occurrence of `r fomoarea3apr['Syringodium filiforme']`\%, July frequency occurrence of `r fomoarea3jul['Syringodium filiforme']`\%).  Changes in macroalgae frequency occurrence were much more pronounced than for seagrasses across months and by area. Red macroalgae was the dominant group across all months and areas, with the highest frequency occurrences observed in April (`r fomoarea1apr['Red']`\% in area 1, `r fomoarea3apr['Red']`\% in area 3). Cyanobacteria frequency occurrence peaked in June, with greater coverage in area 3 (`r fomoarea3jun['Cyanobacteria']`\%) compared to area 1 (`r fomoarea1jun['Cyanobacteria']`\%).  Green macroalgae had the second lowest frequency occurrence, followed by Brown macroalgae that was only observed at one transect.

Statistical comparisons of results in Figure \@ref(fig:trnfrq) shown in Tables \@ref(tab:mcrtab) and \@ref(tab:savtab) further suggested that seagrass trends were minor compared to those of macroalgae.  Results by month showed that seagrass frequency occurrence, by species and area, had insignificant changes ($p > 0.05$) over the study period.  Although trends were not significant, estimated medians in each month in Table \@ref(tab:savtab) showed similar changes for seagrasses as in Figure \@ref(fig:trnex), with generally higher coverage estimates in April.  These changes were more pronounced in area 3 as compared to area 1.  For macroalgae (Table \@ref(tab:mcrtab)), significant differences between months were observed for red and cyanobacteria macroalgae in both areas 1 and 3, although cyanobacteria differences for area 3 were only marginally significant ($p = `r cyanoarea3`$).  Coverage of cyanobacteria in area 1 during June was significantly higher than coverage in April.  No significant differences between months were observed for the green macroalgae group.

@@

```{r trnfrq, fig.cap = paste0('Frequency occurrence estimates for (a) area 1 and (b) area 3 (Figure \\@ref(fig:map)a) for seagrass (top) and macroalgae (bottom) rapid response transect surveys across all transects (n = ', ntrnscnt, ') near Piney Point.  Estimates are grouped by sample months in 2021. Frequency occurrences are absolute for each taxa based on presence/absence, whereas the total frequency occurrence applies to any taxa observed on each transect.')}
knitr::include_graphics(here('figs/trnfrq.jpeg'))
```

@@

```{r mcrtab}
cap.val <- 'Comparison of macroalgae frequency occurrence by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of frequency occurrence between months for macroalgae groups and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly frequency occurrences for each macroalgae group in each area.  Rows that share a letter within each area and macroalgae group combination have frequency occurrences that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- mcrtab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```

```{r savtab}
cap.val <- 'Comparison of seagrass species frequency occurrence by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of frequency occurrence between months for seagrass species and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly frequency occurrences for each seagrass species in each area.  Rows that share a letter within each area and seagrass species combination have frequency occurrences that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- savtab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```

>>>>

load(file = here('tables/wqsumtab.RData'))
load(file = here('tables/stktab.RData'))
load(file = here('tables/wqcmptab.RData'))
load(file = here('tables/savfotab.RData'))
load(file = here('tables/mcrfotab.RData'))

@@

Differences in observations between months for water quality, seagrass, and macroalgae within each area (Figure \@ref(fig:map)a) were evaluated using a Kruskal-Wallis one-way analysis of variance (ANOVA) followed by multiple comparisons using 2-sided Mann-Whitney U tests [@Hollander13]. Probability values were adjusted using the sequential Bonferroni method described in [@Holm79] to account for the increased probability of Type I error rates with multiple comparisons. An adjusted p-value < 5\% ($\alpha$ = 0.05) was considered a significant difference between months.  For water quality variables, long-term monthly averages from long-term monitoring data were subtracted from 2021 observations to account for normal seasonal variation not attributed to potential effects from Piney Point.  Similar corrections were not done for monthly comparisons of seagrass and macroalgae data because comparable long-term seasonal data do not exist. When appropriate, methods were used to accommodate measured concentrations in water quality variables that were below detection.  These included summary statistics (e.g., median, mean, and standard deviation) following estimates of the empirical cumulative distribution functions for each parameter using the Kaplan-Meier method for censored data [@Helsel05;@Lee20].

@@

From April through September 2021, `r nwq` samples were collected for chl-a, dissolved oxygen, total nitrogen, total phosphorus, total ammonia nitrogen, nitrate/nitrite, pH, salinity, secchi depth, and temperature (Table \@ref(tab:wqsumtab)).  Of these samples, `r outrng`\% were outside of the normal range defined by the long-term monthly monitoring data for the baseline period from 2006 to 2020 (below for Secchi depth, above for all others). The percentage of observations outside of the normal range varied by location and parameter.  For chl-a, `r chlabv[['1']]`\% of the observations were above the normal range for area 1, whereas only `r chlabv[[2]]`\% and `r chlabv[[3]]`\% were above for areas 2 and 3, respectively. Total nitrogen concentrations were above the normal range for `r tnabv[['1']]`\% of observations in area 1, whereas concentrations were above for `r tnabv[['2']]`\% of observations in area 2 and `r tnabv[['3']]`\% in area 3.  Secchi observations were below the normal range for `r secblw[['1']]`\% of observations in area 1 and for `r secblw[['2']]`\% and `r secblw[['3']]`\% of observations in areas 2 and 3.  Notable differences were also observed for dissolved oxygen (e.g., `r dosabv[['1']]`\% were above in area 1, `r dosabv[['2']]`\% in area 2).  Physical parameters (salinity, temperature) were generally within range over the study period. Inorganic nitrogen (ammonia, nitrate/nitrite) was generally within range, although initial time series showed much higher concentrations for ammonia in April near area 1, similar to the discharge measurements in Table \@ref(tab:stktab). Spatial variation among the parameters showed that values were generally above the normal range (or below for Secchi depth) for many locations near Piney Point, Anna Maria Sound, and the northern mouth of Tampa Bay (Figure \@ref(fig:wqmap)).

Total nitrogen, chl-a, and secchi depth followed a temporal progression in 2021 that was distinct from the long-term seasonal trend estimated from historical data (Figure \@ref(fig:wqgamtrnds)).  For area 1, total nitrogen and chlorophyll concentrations were frequently above normal ranges during April.  Concentrations decreased slightly until June and July when values increased again above the seasonal expectation. Many secchi observations in area 1 were lower than normal in April and July.  Observations in areas 2 and 3 were more often within the normal seasonal range, with some exceptions for total nitrogen and chl-a in area 3 in  April, May, and July.  Statistical comparisons between months for total nitrogen, chl-a, and secchi depth (Table \@ref(tab:wqcmptab)) supported the results in Figure \@ref(fig:wqgamtrnds). Kruskal-Wallis tests that assessed if at least one of the months had significantly different observations for each parameter were significant for total nitrogen, chl-a, and secchi depth for area 1 and for total nitrogen and chl-a for areas 2 and 3 (Table \@ref(tab:wqcmptab)). Results of multiple comparison tests that evaluated differences between pairs of months generally showed that April/May were different from June/July depending on area and parameter.  Observations in the later months were generally higher (or lower for Secchi) corresponding to increasing *K. brevis* abundances.

@@

# fo by month, area
fomo <- trnsum %>% 
  ungroup() %>% 
  mutate(
    foest = round(100 * foest, 1), 
    mo = month(date)
  )
fomoarea1 <- fomo %>% 
  filter(area == 'Area 1')
fomoarea3 <- fomo %>% 
  filter(area == 'Area 3')
fomoarea1sg <- fomoarea1 %>% 
  group_by(taxa) %>% 
  summarise(foest = mean(foest)) %>% 
  deframe

fomoarea1apr <- fomoarea1 %>% 
  filter(mo %in% 4) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea1jul <- fomoarea1 %>% 
  filter(mo %in% 7) %>% 
  select(taxa,foest) %>% 
  deframe
fomoarea3apr <- fomoarea3 %>% 
  filter(mo %in% 4) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea3jul <- fomoarea3 %>% 
  filter(mo %in% 7) %>% 
  select(taxa,foest) %>% 
  deframe
fomoarea1jun <- fomoarea1 %>% 
  filter(mo %in% 6) %>% 
  select(taxa, foest) %>% 
  deframe
fomoarea3jun <- fomoarea3 %>% 
  filter(mo %in% 6) %>% 
  select(taxa, foest) %>% 
  deframe

# kw pvals
cyanoarea3 <- mcrfotab %>% 
  mutate(
    Area = ifelse(Area == '', NA, Area)
  ) %>% 
  fill(Area) %>% 
  filter(Area == 3 & `Macroalgae group` == 'Cyanobacteria') %>% 
  pull(pval) %>% 
  as.numeric %>% 
  round(3)
```

A total of `r ntrns` transects were sampled for seagrasses and macroalgae from April through September, each visited on average `r avevis` times per month. Turtle grass (*T. testudinum*) was the dominant seagrass species with frequency occurrence of `r foall['Thalassia testudinum']`\% across all locations and sample dates.  Manatee grass (*S. filiforme*) and shoal grass (*H. wrightii*) had similar coverage across all transects, with frequency occurrences of `r foall['Syringodium filiforme']`\% and `r foall['Halodule wrightii']`\%, respectively. The frequency occurrences of seagrasses near Piney Point were similar to the long-term record of seagrass transect data available for Tampa Bay [@Sherwood17].  At the baywide scale, shoal grass is the dominant species, whereas turtle grass is more common in euhaline waters closer to the Gulf.  Macroalgae observed along the transects also varied in coverage, with red macroalgae groups having the highest frequency occurrence of `r foall['Red']`\%.  Common taxa in the red group included *Gracilaria spp.* and *Acanthophora sp.*.  Green and cyanobacteria macroalgae were less common, with frequency occurrences of `r foall['Green']`\% and `r foall['Cyanobacteria']`\%.  Common taxa in the green group included *Ulva spp.* and *Caulerpa spp.*, whereas *Dapis spp.* was the only cyanobacteria macroalgae observed.  Brown macroalgae in the genus *Feldmannia* were only observed at `r foall['Brown']`\% frequency occurrence.

Temporal progression of seagrasses and macroalgae varied across the months, although a typical pattern observed at many of the transects is shown in Figure \@ref(fig:trnex).  Transect S3T6 is located less than one kilometer to the north of Port Manatee.  The site is dominated by manatee grass that was observed at nearly all of the sample points along the transect at varying coverages.  Overall abundance of seagrass did not change dramatically from April 22nd through September.  However, macroalgal abundances changed over the course of sampling similar to many of the other transects sampled during the study.  Red macroalgae were present in high abundances from April to May.  *Dapis sp.* was first observed on May 24th and was present at all of the sample locations on June 4th and 15th.  *Dapis sp.* persisted through June and July, but not was observed after July 20th.  Green macroalgae taxa were first observed in July, although at generally low abundances. Red macroalgae were the dominant taxa at the end of September.

Monthly summaries in frequency occurrence by area (Figure \@ref(fig:trnfrq)) provided an indication of seagrass and macroalgae trends across all transects.  No transects were sampled in area 2 to the north of Piney Point and no transects were sampled past September in Area 1.  Area 1 had generally stable frequency occurrence over time of the three primary seagrass species observed in the bay, with turtle grass (*T. testudinum*) occurring slightly more often (average frequency occurrence across months `r fomoarea1sg['Thalassia testudinum']`\%) than shoal grass (*H. wrightii*, `r fomoarea1sg['Halodule wrightii']`\%)) or manatee grass (*S. filiforme*, `r fomoarea1sg['Syringodium filiforme']`\%)).

, although a slight decrease over time in shoal grass (*H. wrightii*) was observed (April frequency occurrence of `r fomoarea1apr['Halodule wrightii']`\%, July frequency occurrence of `r fomoarea1jul['Halodule wrightii']`\%).  A similar decline in coverage of shoal grass was also observed in area 3 (April frequency occurrence of `r fomoarea3apr['Halodule wrightii']`\%, July frequency occurrence of `r fomoarea3jul['Halodule wrightii']`\%). Frequency occurrence of manatee grass (*S. syringodium*) also declined slightly from April to July in area 3 (April frequency occurrence of `r fomoarea3apr['Syringodium filiforme']`\%, July frequency occurrence of `r fomoarea3jul['Syringodium filiforme']`\%).

Changes in macroalgae frequency occurrence were much more pronounced than for seagrasses across months and by area. Red macroalgae was the dominant group across all months and areas, with the highest frequency occurrences observed in April (`r fomoarea1apr['Red']`\% in area 1, `r fomoarea3apr['Red']`\% in area 3). Cyanobacteria frequency occurrence peaked in June, with greater coverage in area 3 (`r fomoarea3jun['Cyanobacteria']`\%) compared to area 1 (`r fomoarea1jun['Cyanobacteria']`\%).  Green macroalgae had the second lowest frequency occurrence, followed by Brown macroalgae that was only observed at one transect.

Statistical comparisons of results in Figure \@ref(fig:trnfrq) shown in Tables \@ref(tab:mcrfotab) and \@ref(tab:savfotab) further suggested that seagrass trends were minor compared to those of macroalgae.  Results by month showed that seagrass frequency occurrence, by species and area, had insignificant changes ($p > 0.05$) over the study period.  Although trends were not significant, estimated medians in each month in Table \@ref(tab:savfotab) showed similar changes for seagrasses as in Figure \@ref(fig:trnex), with generally higher coverage estimates in April.  These changes were more pronounced in area 3 as compared to area 1.  For macroalgae (Table \@ref(tab:mcrfotab)), significant differences between months were observed for red and cyanobacteria macroalgae in both areas 1 and 3, although cyanobacteria differences for area 3 were only marginally significant ($p = `r cyanoarea3`$).  Coverage of cyanobacteria in area 1 during June was significantly higher than coverage in April.  No significant differences between months were observed for the green macroalgae group.

@@

```{r trnfrq, fig.cap = paste0('Frequency occurrence estimates for (a) area 1 and (b) area 3 (Figure \\@ref(fig:map)a) for macroalgae (top) and seagrass (bottom) rapid response transect surveys across all transects (n = ', ntrnscnt, ') near Piney Point.  Estimates are grouped by sample months in 2021. Frequency occurrences are absolute for each taxa based on presence/absence, whereas the total frequency occurrence applies to any taxa observed on each transect. Points are offset slightly for readability.')}
knitr::include_graphics(here('figs/trnfrq.jpeg'))
```

@@

```{r mcrfotab}
cap.val <- 'Comparison of macroalgae frequency occurrence by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of frequency occurrence between months for macroalgae groups and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly frequency occurrences for each macroalgae group in each area.  Rows that share a letter within each area and macroalgae group combination have frequency occurrences that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- mcrfotab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```

```{r savfotab}
cap.val <- 'Comparison of seagrass species frequency occurrence by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of frequency occurrence between months for seagrass species and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly frequency occurrences for each seagrass species in each area.  Rows that share a letter within each area and seagrass species combination have frequency occurrences that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- savfotab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```

### 2021-10-20 supplement.Rmd
added sg and macro abundance tables and figures

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(fig.path = 'figs/', warning = F, message = F, echo = F, cache = F, dev.args = list(family = 'serif'), dpi = 300, warning = F,
  fig.process = function(x) {
  x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  if (file.rename(x, x2)) x2 else x
  })
```

@@

@@

>>>>

```{r setup, echo = F, warning = F, message = F, results = 'hide'}
# figure path, chunk options
knitr::opts_chunk$set(fig.path = 'figs/', warning = F, message = F, echo = F, cache = F, dev.args = list(family = 'serif'), dpi = 300, warning = F,
  fig.process = function(x) {
  x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
  if (file.rename(x, x2)) x2 else x
  })

box::use(
  here[here],
  dplyr[...]
)

load(file = here('tables/mcrabutab.RData'))
load(file = here('tables/savabutab.RData'))
```

# Figures

@@

```{r trnabu, fig.cap = 'Abundance estimates (+/- 95% confidence) for (a) area 1 and (b) area 3 (Figure 1a) for macroalgae (top) and seagrass (bottom) rapid response transect surveys across all transects (n = 38) near Piney Point.  Estimates are grouped by sample months in 2021. Points are offset slightly for readability.'}
knitr::include_graphics(here('figs/trnabu.jpeg'))
```

@@

# Tables

```{r mcrabutab}
cap.val <- 'Comparison of macroalgae Braun-Blanquet abundances by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of abundances between months for macroalgae groups and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly abundances for each macroalgae group in each area.  Rows that share a letter within each area and macroalgae group combination have abundances that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- mcrabutab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```

```{r savabutab}
cap.val <- 'Comparison of seagrass species Braun-Blanquet abundances by areas of interest (Figure \\@ref(fig:map)a) and month.  Overall signifance of differences of abundances between months for seagrass species and area combination are shown with Chi-squared statistics based on Kruskall-Wallis rank sum tests.  Multiple comparisons with Mann-Whitney U tests (Comp. column) were used to evaluate pairwise monthly abudances for each seagrass species in each area.  Rows that share a letter within each area and seagrass species combination have abundances that are not significantly different between month pairs. Probability values were adjusted for the pairwise comparisons using the Bonferroni method in @Holm79. ** p < 0.005, * p < 0.05, blank is not significant at $\\alpha$ = 0.05.'
totab <- savabutab %>% 
  select(-pval)
knitr::kable(totab, booktabs = T, caption = cap.val)
```
